### 2019-01-31 metrics3.tex
Metrics

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{lnahe}
	\caption{Summary of the regression}\label{fig1}
\end{figure}

\begin{enumerate}[(a)]
	\item First of all, all three coefficients are significant (on 5\% significance level). Of course it would be more correct to use coeftest() in order to get heteroskedasticity-robust standard errors, but p-value is such a small that this procedure do not change our conclusion about significance (only standard errors and t-statistics change slightly). Then, using logic similar to the Problem 1, we can infer that, other things equal each additional year of education increases the average per hour salary by 8\%, holding other things equal each additional year of life increases per hour earnings by 0.6\%, and finally, holding other things equal women get on average 22\% less per hour salary than men.
	\item RMSE is approximately equal to $0.465$.
	\item Confidence interval for the gender gap in earnings is
	\begin{align*}
	(-0.254035195; -0.194271848)
	\end{align*}
	i.e. with probability 0.95 the interval $(19\%; 25\%)$ cover the value which indicates by how much percent the average per hour salary for women is less than the one for men.
	\item Based on the summary of regression, the marginal value of the high school diploma is 8.3, i.e. holding gender and age fixed, receiving a high school diploma after 12 years of studying against 12 years without diploma increases average per hour earnings by 8.3\%. Using coeftest() we can conclude that on the 5\% significance level we reject the hypothesis that the marginal value of high school diploma is zero. 8.3\% is not a huge value, in fact, I suppose average the salary of people with high school diploma much more higher than those without diploma.
	\item 
	Estimating the following regression model
	\begin{align*}
	\ln(\text{ahe}) = \beta_0 + \beta_1 \text{hsdipl} + \beta_2 \text{yrseduc} + \beta_3 \text{age} + \beta_4\text{female} + u
	\end{align*}
	one can get:
	\begin{align*}
	\ln \text{ahe}_{12} - \ln \text{ahe}_{10} &= \beta_1 + 2 \beta_2\\
	\frac{\text{ahe}_{12} - \text{ahe}_{10}}{\text{ahe}_{10}} &= V = e^{\beta_1 + 2\beta_2} - 1 \\
	\hat{V} = e^{\hat{\beta_1} + 2\hat{\beta_2}} - 1 &\approx 0.285
\end{align*}
I think it is a rather reasonable estimate.
\item 
\begin{align*}
\prob (\alpha_1 < e^{\hat{\beta_1} + 2 \hat{\beta_2}} - 1 < \alpha_{2}) = 0.95\\
\prob (\ln(\alpha_1 + 1) < \hat{\beta_1} + 2\hat{\beta_2} < \ln(\alpha_2 +1)) = 0.95
\end{align*}
Thus, we need to compute a confidence interval for $\hat{\beta_1} + 2 \hat{\beta_2}$. To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on $1$, \textit{hsdipl } $+ 1$, \textit{yrseduc } $+2$, \textit{age} and \textit{female}.
\end{enumerate}
\end{document}

>>>>

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{lnahe}
	\caption{Summary of regression}\label{fig1}
\end{figure}

\begin{enumerate}[(a)]
	\item First of all, all three coefficients are significant (on 5\% significance level). Of course it would be more correct to use coeftest() in order to get heteroskedasticity-robust standard errors, but p-value is such a small that this procedure do not change our conclusion about significance (only standard errors and t-statistics change slightly). Then, using logic similar to the Problem 1, we can infer that, other things equal each additional year of education increases the average per hour salary by 8\%, holding other things equal each additional year of life increases per hour earnings by 0.6\%, and finally, holding other things equal women get on average 22\% less per hour salary than men.
	\item RMSE is approximately equal to $0.465$.
	\item Confidence interval for the gender gap in earnings is
	\begin{align*}
	(-0.254035195; -0.194271848)
	\end{align*}
	i.e. with probability 0.95 the interval $(19\%; 25\%)$ cover the value which indicates by how much percent the average per hour salary for women is less than the one for men.
	\item Based on the summary of regression, the marginal value of the high school diploma is 8.3, i.e. holding gender and age fixed, receiving a high school diploma after 12 years of studying against 12 years without diploma increases average per hour earnings by 8.3\%. Using coeftest() we can conclude that on the 5\% significance level we reject the hypothesis that the marginal value of high school diploma is zero. 8.3\% is not a huge value, in fact, I suppose average the salary of people with high school diploma much more higher than those without diploma.
	\item 
	Estimating the following regression model
	\begin{align*}
	\ln(\text{ahe}) = \beta_0 + \beta_1 \text{hsdipl} + \beta_2 \text{yrseduc} + \beta_3 \text{age} + \beta_4\text{female} + u
	\end{align*}
	one can get:
	\begin{align*}
	\ln \text{ahe}_{12} - \ln \text{ahe}_{10} &= \beta_1 + 2 \beta_2\\
	\frac{\text{ahe}_{12} - \text{ahe}_{10}}{\text{ahe}_{10}} &= V = e^{\beta_1 + 2\beta_2} - 1 \\
	\hat{V} = e^{\hat{\beta_1} + 2\hat{\beta_2}} - 1 &\approx 0.285
\end{align*}
I think it is a rather reasonable estimate.
\item 
\begin{align*}
\prob (\alpha_1 < e^{\hat{\beta_1} + 2 \hat{\beta_2}} - 1 < \alpha_{2}) = 0.95\\
\prob (\ln(\alpha_1 + 1) < \hat{\beta_1} + 2\hat{\beta_2} < \ln(\alpha_2 +1)) = 0.95
\end{align*}
Thus, we need to compute a confidence interval for $\hat{\beta_1} + 2 \hat{\beta_2}$. To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on \textit{hsdipl } $- 1$, \textit{yrseduc } $-2$, \textit{age} and \textit{female}. It is easy to observe that in that case the intercept will be $\beta_0 + \beta_1 + 2\beta_2$. Using confint() for this new intercept we can get the following:
\begin{align*}
\prob (1.375500454 &< \widehat{\beta_0 + \beta_1 + 2\beta_2} < 1.585969536) = 0.95\\
\prob(0.1458 &< \hat{\beta_1} + 2\hat{\beta_2} < 0.3563) = 0.95
\end{align*}
Thus, $\alpha_1 = 0.157$, $\alpha_2 = 0.428$ and
\begin{align*}
\prob(0.157 < \hat{V} < 0.428) = 0.95
\end{align*}
\end{enumerate}
\section*{Problem 4}
The summary of regression is depicted below (see Fig. \ref{fig2})

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{square}
	\caption{Summary of the regression}\label{fig2}
\end{figure}

\begin{enumerate}[(a)]
	\item Sign of the coefficient on \textit{age} is +, while on \textit{age}$^2$ is -. Denote $y = \ln \text{ahe}, x_1 = \text{ age}, x_3 = \text{ female}, x_4 = \text{hsdipl}$. Since 
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
	that means that
	\begin{align*}
	\frac{d \ln y}{d x_1} = \beta_1 + 2\beta_2x_1\\
	\frac{dy}{y} = \beta_1dx_1 + 2\beta_2 x_1 dx_1 \\
	\frac{\Delta y}{y} \approx \beta_1\Delta x_1 + 2\beta_2 x_1 \Delta x_1
	\end{align*}
	Since $\beta_1 > 0$ while $\beta_2 < 0$ that means that: the older the person becomes - the less (on average) increase in salary each additional year brings. For example, if $x_1 = 20, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 2.96 \%$, but if $x_1 = 40, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 0.86 \%$.
	\item The relation is linear $\iff$ $\beta_2 = 0$. Using coeftest() we reject this hypothesis on the 5\% significance level, i.e. the relation is quadratic.
	\item The estimated gender gap has changed from -0.224 to -0.225, which is not a significant change in a real-world sense. Econometric reason is that first of all $\text{age}^2$ and $\text{female}$ are almost uncorrelated (sample correaltion coeeficient is about 0.007) and moreover slope coefficient on $\text{age}^2$ is also very small, as a result including $\text{age}^2$ barely changes the estimated gender gap.
	\item Let us estimate the following regression
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_{23}x_1^3+ \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
	the relation is cubic $\iff \beta_{23} = 0$. Using coeftest() we can conclude that on the 5\% significance level we cannot reject this hypothesis, i.e. the relation is quadratic.
	\item Here we need to test the hypothesis $\beta_2 = 0, \beta_{23} = 0$. The output of linearHypothesis() is depicted below (see Fig. \ref{fig3}).

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.8\textwidth]{hyptest}
		\caption{Hypothesis test}\label{fig3}
	\end{figure}

	Thus, the hypothesis can be rejected on the 5\% significance level. That is the relation is more likely a polynomial of up to cubic degree. I recomend the specification
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
\end{enumerate}
\end{document}
